Formalizing Stack Safety as a Security Property

### 2022-11-28 submission5/noexceptions.tex
Almost there

In addition to memory and registers, a machine has a program counter \(\PCname\)
and may have additional architecture-specific state; particularly relevant is state
that is used for security enforcement. We treat this as a set \(\mathcal{P}\) with
\(\PCname \in \mathcal{P}\).

@@

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS + \mathcal{P}\). A {\em machine state}
\(\mach\) is a map from components to words, written \(\mach[\component]\).
\apt{As we discussed, this seems too restrictive a definition for $\mathcal{P}$.
  For example, for tagged machines it might contain a mapping from $\WORDS + \REGS$ to tags.}

\sna{It occurs to me, based on some of your comments further down, that the old machine
  states were unclear. My intention was not for \(\mach\) to be a map, but for
  it to be an arbitrary type on which \(\cdot[\cdot]\) was defined. (Akin to
  the coq idiom of a record with a field t:Type and proj:t\(\rightarrow\)k\(\rightarrow\)w.)
  Therefore, implicitly, there could be additional state for enforcement and other
  purposes. The new presentation with P makes that explicit but is less elegant.
}

@@

\sna{And the payoff of \(\mathcal{P}\) being explicit: we can just state this.}

@@

To characterize this situation, we define a state element to be \emph{irrelevant}
if its value can be replaced with any other value without changing the observable behavior
of the machine. More precisely, for a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\)
are {\em \(\components\)-variants} if they agree on the value of every
element except those in \(\components\). \apt{Explicitly note that this excludes elements of
  of $\mathcal{P}$ (assuming that is indeed the case).} The elements of \(\components\) are \emph{irrelevant}
in \(\mach\) if every variant state produces an equivalent observation trace
(see Section \ref{sec:events}); all other elements are \emph{relevant}.
We characterize \emph{observational integrity} as the property that
every relevant element that is \(\sealed\) under the callee's view is restored
to its original value at the return point.

@@

\paragraph*{Moving Forward}\apt{Suggest omit this para.}

In the next section we will define traces and their (hyper-)properties.
Then in Section \ref{sec:facts} we will define a mechanism to make assertions
about what will happen on a future return, and Section \ref{sec:props} will use these
to give the formal definitions of integrity and confidentiality.

@@

\section{Enforcement}

\paragraph{Protecting Enforcement State}

>>>>

In addition to memory and registers, a machine has a program counter \(\PCname\)
and may have additional architecture-specific state; particularly relevant is state
that is used for security enforcement. We treat this as a set \(\mathcal{P}\) with
\(\PCname \in \mathcal{P}\). Collectively these are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS | \mathcal{P}\).

@@

A machine has a set \(\mathcal{V}\) of values such that \(\WORDS \subseteq \mathcal{V}\).
A {\em machine state} \(\mach\) is a map from components to values, written \(\mach[\component]\).

@@

@@

To characterize this situation, we define a state element to be \emph{irrelevant}
if its value can be replaced with any other value without changing the observable behavior
of the machine. More precisely, for a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if they
agree on the value of every element except those in \(\components\). This includes
public elements such as those in \(\mathcal{P}\).
The elements of \(\components\) are \emph{irrelevant}
in \(\mach\) if every variant state produces an equivalent observation trace
(see Section \ref{sec:events}); all other elements are \emph{relevant}.
We characterize \emph{observational integrity} as the property that
every relevant element that is \(\sealed\) under the callee's view is restored
to its original value at the return point.

@@

@@

\section{Enforcement and Testing}

We use our properties to test the stack safety micro-policies of Roessler and DeHon,
particularly their Lazy Tagging and Clearing. Our underlying machine is a RISC-V
based on the MIT RISC-V formalism \cite{} and extended with {\it metadata tags},
represented as additional state in \(\mathcal{P}\), one for each register, address,
and the program counter:
\[\mathcal{P} ::= t_\reg | t_\addr | t_\PCname\]
%
We divide values into normal memory words and tags drawn from a policy-specific
set of tags \(\mathcal{T}\), \(\mathcal{V} ::= \mathcal{W} | \mathcal{T}\).
The machine's step function corresponds to that of the basic RISC-V machine,
with the behavior of the tags at each step parameterized by a {\it micro-policy},
which may turn a normal step into an error (represented by returning to the same state.)

\paragraph*{Protecting Enforcement State}
